Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-01 docs/evol.tex
new style

\documentclass[conference]{IEEEtran}

@@

\author{
\IEEEauthorblockN{Dimitris Mitropoulos\IEEEauthorrefmark{1}, Vassilios Karakoidas\IEEEauthorrefmark{1},
Georgios Gousios\IEEEauthorrefmark{2}, Panagiotis Louridas\IEEEauthorrefmark{1},
Diomidis Spinellis\IEEEauthorrefmark{1}}\\
\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Management Science and Technology\\
Athens University of Economics and Business\\
\{dimitro, bkarak, louridas, dds\}@aueb.gr}\\
\IEEEauthorblockA{\IEEEauthorrefmark{2}
Software Engineering Research Group\\
Delft University of Technology\\
G.Gousios@tudelft.nl}}

@@

@@

\begin{IEEEkeywords}
Software Bugs, Static Analysis, Software Evolution, Software
Security, Maven, FindBugs.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

@@

A security bug is a programming error that introduces a potentially
exploitable weakness into a computer system~\cite{SSL12, TJBD11}. This weakness could lead to a
security breach with unfortunate consequences in different layers, like databases,
native code, applications, libraries and others. Despite the significant
effort to detect and eliminate such bugs~\cite{SZ12}, little attention has been paid to
study them in relation to software evolution~\cite{L96, LRWPT97, IB06, RGMA06}.
In this paper we present how we used a large software ecosystem to analyze
how evolving software packages are related to the different types of security
vulnerabilities.

One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it. For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10, HP07, HP04, HW08}. Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the Maven\footnote{\url{http://maven.apache.org/}}
repository (approximately 265GB of data). Then we observed the changes that
involved the security bugs and their characteristics. This research builds upon
our earlier work on the topic~\cite{MGS12}.

@@

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistency of a bug. In essence, to check if specific bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the issue~\cite{BP84, SYTP85,
NBZ06, GKMS00}. Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the oportunity to see if a project version that is a dependency to
a large number of others contains a high rate of security bugs.
In this case we can strenghten the relationship between Linus' Law and
software security as Meneely et al. have already indicated~\cite{MW10}.

@@

To see whether bugs flock together we performed pairwise correlations
between all bug categories. We calculated the correlations between the
number of distinct bugs that appeared in a project throughout its
lifetime. The results can be seen in Table~\ref{tbl:corrmatrix} and
Figure~\ref{fig:corrplot}. We found significant correlations between
all pairs apart from two pairs involving the {\it Security High}
category. In general, the {\it Security High} category showed the
weekest correlations with the other categories. Our results show that
in general bugs do flock together. We do not find projects with only a
certain kind of bug; bugs come upon projects in swarms of different
kinds. Bugs of the {\it Security High} category, though, are
different: they are either not associated with other bugs, or only
weakly so. Perhaps it takes a special kind of blunder to make it a
security hazard.
\begin{table}
  \begin{threeparttable}
    \centering
    \caption{Correlation matrix}
    \label{tbl:corrmatrix}
    \input{corrmatrix.tex}
    \begin{tablenotes}
      \footnotesize
      \item The matrix presents pairwise correlations between the
        following bug categories in order: Security High, Security
        Low, Style, Bad Practice, Correctness, MT Correctness,
        Performance, i18n, Experimental.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

{\it Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11, B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instanse, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.

@@

\bibliographystyle{IEEEtran}
\bibliography{evol}

>>>>

\documentclass[conference]{llncs}

@@

@@

\author{Dimitris Mitropoulos    \and
       Vassilios Karakoidas	\and
       Georgios Gousios         \and
       Panos Louridas		\and
       Diomidis Spinellis
}
\institute{
       	Department of Management Science and Technology\\
	Athens University of Economics and Business\\
      	\email{\{dimitro, bkarak, louridas, dds\}@aueb.gr}\\
	Software Engineering Research Group\\
	Delft University of Technology\\
	\email{G.Gousios@tudelft.nl}
}

@@

\begin{keywords}
Software Bugs, Static Analysis, Software Evolution, Software
Security, Maven, FindBugs.
\end{keywords}

@@

A security bug is a programming error that introduces a potentially
exploitable weakness into a computer system~\cite{SSL12}\cite{TJBD11}. This weakness could lead to a
security breach with unfortunate consequences in different layers, like databases,
native code, applications, libraries and others. Despite the significant
effort to detect and eliminate such bugs~\cite{SZ12}, little attention has been paid to
study them in relation to software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}.
In this paper we present how we used a large software ecosystem to analyze
how evolving software packages are related to the different types of security
vulnerabilities.

One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it. For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10}\cite{HP07}\cite{HP04}\cite{HW08}. Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the Maven\footnote{\url{http://maven.apache.org/}}
repository (approximately 265GB of data). Then we observed the changes that
involved the security bugs and their characteristics. This research builds upon
our earlier work on the topic~\cite{MGS12}.

@@

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistency of a bug. In essence, to check if specific bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
did) but also with other bug categories. In the same manner, we tried to check
the relation between a version's size and its' security bugs knowing the
that research has been contradictory on the the issue~\cite{BP84}\cite{SYTP85},
\cite{NBZ06}\cite{GKMS00}. Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure, gave us the oportunity to see if a project version that is a dependency to
a large number of others contains a high rate of security bugs.
In this case we can strenghten the relationship between Linus' Law and
software security as Meneely et al. have already indicated~\cite{MW10}.

@@

To see whether bugs flock together we performed pairwise correlations
between all bug categories. We calculated the correlations between the
number of distinct bugs that appeared in a project throughout its
lifetime. The results can be seen in Table~\ref{tbl:corrmatrix} andsplncs.bst
Figure~\ref{fig:corrplot}. We found significant correlations between
all pairs apart from two pairs involving the {\it Security High}
category. In general, the {\it Security High} category showed the
weekest correlations with the other categories. Our results show that
in general bugs do flock together. We do not find projects with only a
certain kind of bug; bugs come upon projects in swarms of different
kinds. Bugs of the {\it Security High} category, though, are
different: they are either not associated with other bugs, or only
weakly so. Perhaps it takes a special kind of blunder to make it a
security hazard.
\begin{table}
  \begin{threeparttable}
    \centering
    \caption{Correlation matrix}
    \label{tbl:corrmatrix}
    \input{corrmatrix.tex}
    \begin{tablenotes}
      \footnotesize
      \item The matrix presents pairwise correlations between the
        following bug categories in order: Security High, Security
        Low, Style, Bad Practice, Correctness, MT Correctness,
        Performance, i18n, Experimental.
    \end{tablenotes}
  \end{threeparttable}
\end{table}

@@

{\it Querying techniques} are used to answer a broad range of questions
regarding the evolution history of a project~\cite{HG05}. Bhattacharya et
al.~\cite{BN11}\cite{B11} proposed a framework that is based upon
recursively enumerable languages. The framework can correlate software
bugs with developers in various ways. For instanse, return the list of
bugs fixed by a specific developer. Fischer et al.~\cite{FPG03} proposed
an approach for populating a release history database that combines code
information with bug tracking data. In this way, a developer can couple files
that contain common bugs, estimate code maturity with respect to the bugs
and others. The ``Ultimate Debian Database''~\cite{NZ10} is an {\sc sql}-based
framework that integrates information about the Debian project from various
sources to answer queries related to software bugs and source code.

@@

\bibliographystyle{splncs}
\bibliography{evol}
